PANCHANGE_data_preprocessing

### 2020-11-16 PANCHANGE_data_merging.Rmd
education, employment & employment change sensitivity

# Import data
```{r export data}
glad <- read_rds(path = paste0(data_path, "/glad.rds"))
dim(glad)

edgi <- read_rds(path = paste0(data_path, "/edgi.rds"))
dim(edgi)

nbr <- read_rds(path = paste0(data_path, "/nbr.rds"))
dim(nbr)

ramp <- read_rds(path = paste0(data_path, "/ramp.rds"))
dim(ramp)
```

>>>>

# Import data
```{r export data}
glad <- read_rds(file = paste0(data_path, "/glad.rds"))
dim(glad)

edgi <- read_rds(file = paste0(data_path, "/edgi.rds"))
dim(edgi)

nbr <- read_rds(file = paste0(data_path, "/nbr.rds"))
dim(nbr)

ramp <- read_rds(file = paste0(data_path, "/ramp.rds"))
dim(ramp)
```

### 2020-11-16 PANCHANGE_data_preprocessing2020-07-15.Rmd
Knitted exclusion descriptives script
education, employment & employment change sensitivity

# Specify cohort
```{r Specify cohort}
GLAD = FALSE
EDGI = FALSE
NBR = TRUE
RAMP = FALSE
GLAD_and_EDGI = FALSE
```

@@

```{r source file path}
#source raw data directory: data.raw_path
source("..//PANCHANGE_raw_path.R")
```

@@

## GLAD and EDGI
### Create data frame 
```{r GLAD and EDGI combined}
glad.and.edgi.dem.raw.id <- rbind(glad.dem.raw.id, edgi.dem.raw.id)
```

@@

glad.employment.raw.id <- glad.employment.raw %>%
  drop_na(externalDataReference) %>% # Drop NAs
  distinct(externalDataReference, .keep_all = TRUE) %>% # Remove duplicates based on ID
  separate(externalDataReference, into = c("Sample", "ID"), sep = 4) %>% # Split ID in Sample and Number
  mutate(
    ID = as.numeric(ID)
    ) %>%
  select(
         Sample, # Sample
         ID, # ID
         employment.phd, # level of education - univeristy
         employment.masters_degree_or_equivalent, # level of education - univeristy
         employment.postgraduate_degree_or_equivalent, # level of education - univeristy
         employment.bachelors_degree_or_equivalent, # level of education - univeristy
         employment.a_levelsas_levels_or_equivalent, # level of education - alevel
         employment.o_levelsgcses_or_equivalent, # level of education - gcse
         employment.cses_or_equivalent, # level of education - cse
         employment.nvq_or_hnd_or_hnc_or_equivalent, # level of education - nvq
         employment.government_work_key_workers, # key worker status
         employment.government_work_key_workers, # key worker status
         employment.unemployed, #key worker routing variable
         employment.zerohours_contract, #key worker routing variable
         employment.stayathome_parent_or_carer, #key worker routing variable
         employment.retired, #key worker routing variable
         employment.student_gcse_or_a_level, #key worker routing variable
         employment.student_university, #key worker routing variable
         employment.became_unemployed, #key worker routing variable
         employment.furloughed_or_paid_leave_government_funded, #key worker routing variable
         employment.furloughed_or_paid_leave_company_funded, #key worker routing variable
         employment.paid_leave_furloughed #key worker routing variable
         ) %>%
  add_numeric(., exclude = exclude_cols_employment) %>%
  mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
  mutate_if(is.numeric, ~na_if(., -99)) %>%
  mutate_if(is.numeric, ~na_if(., -77)) %>%
  mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
  mutate_if(is.factor, ~na_if(., "Don't know")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

@@

edgi.employment.raw.id <- edgi.employment.raw %>%
  drop_na(externalDataReference) %>% # Drop NAs
  distinct(externalDataReference, .keep_all = TRUE) %>% # Remove duplicates based on ID
  separate(externalDataReference, into = c("Sample", "ID"), sep = 4) %>% # Split ID in Sample and Number
#  separate("Login ID", into = c("COPING_Sample", "COPING_ID"), sep = 6) %>% # Split ID in Sample and Number
  mutate(
    ID = as.numeric(ID),
#    COPING_ID = as.numeric(COPING_ID)
    ) %>%
  select(
         Sample, # Sample
         ID, # ID
         COPING_ID = "Login ID",
         employment.phd, # level of education - univeristy
         employment.masters_degree_or_equivalent, # level of education - univeristy
         employment.postgraduate_degree_or_equivalent, # level of education - univeristy
         employment.bachelors_degree_or_equivalent, # level of education - univeristy
         employment.a_levelsas_levels_or_equivalent, # level of education - alevel
         employment.o_levelsgcses_or_equivalent, # level of education - gcse
         employment.cses_or_equivalent, # level of education - cse
         employment.nvq_or_hnd_or_hnc_or_equivalent, # level of education - nvq
         employment.government_work_key_workers, # key worker status
         employment.government_work_key_workers, # key worker status
         employment.unemployed, #key worker routing variable
         employment.zerohours_contract, #key worker routing variable
         employment.stayathome_parent_or_carer, #key worker routing variable
         employment.retired, #key worker routing variable
         employment.student_gcse_or_a_level, #key worker routing variable
         employment.student_university, #key worker routing variable
         employment.became_unemployed, #key worker routing variable
         employment.furloughed_or_paid_leave_government_funded, #key worker routing variable
         employment.furloughed_or_paid_leave_company_funded, #key worker routing variable
         employment.paid_leave_furloughed #key worker routing variable
         ) %>%
  add_numeric(., exclude = exclude_cols_employment_edgi) %>%
  mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
  mutate_if(is.numeric, ~na_if(., -99)) %>%
  mutate_if(is.numeric, ~na_if(., -77)) %>%
  mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
  mutate_if(is.factor, ~na_if(., "Don't know")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

@@

nbr.employment.raw.id <- nbr.employment.raw %>%
  drop_na(subjectid) %>% # Drop NAs
  distinct(subjectid, .keep_all = TRUE) %>% # Remove duplicates based on ID
  separate(subjectid, into = c("Sample", "ID.intm"), sep = 6) %>% # Split ID in Sample and Number
  separate(ID.intm, into = "ID", sep = 7) %>%
  mutate(
    ID = as.numeric(ID)) %>%
  select(
         Sample, # Sample
         ID, # ID
         employment.phd, # level of education - univeristy
         employment.masters_degree_or_equivalent, # level of education - univeristy
         employment.postgraduate_degree_or_equivalent, # level of education - univeristy
         employment.bachelors_degree_or_equivalent, # level of education - univeristy
         employment.a_levelsas_levels_or_equivalent, # level of education - alevel
         employment.o_levelsgcses_or_equivalent, # level of education - gcse
         employment.cses_or_equivalent, # level of education - cse
         employment.nvq_or_hnd_or_hnc_or_equivalent, # level of education - nvq
         employment.government_work_key_workers, # key worker status
         employment.unemployed, #key worker routing variable
         employment.zerohours_contract, #key worker routing variable
         employment.stayathome_parent_or_carer, #key worker routing variable
         employment.retired, #key worker routing variable
         employment.student_gcse_or_a_level, #key worker routing variable
         employment.student_university, #key worker routing variable
         employment.became_unemployed, #key worker routing variable
         employment.furloughed_or_paid_leave_government_funded, #key worker routing variable
         employment.furloughed_or_paid_leave_company_funded, #key worker routing variable
         employment.paid_leave_furloughed #key worker routing variable
         ) %>%
  add_numeric(., exclude = exclude_cols_employment) %>%
  mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
  mutate_if(is.numeric, ~na_if(., -99)) %>%
  mutate_if(is.numeric, ~na_if(., -77)) %>%
  mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
  mutate_if(is.factor, ~na_if(., "Don't know")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

@@

ramp.employment.raw.id <- ramp.employment.raw %>%
  drop_na(`Login ID`) %>% # Drop NAs
  distinct(`Login ID`, .keep_all = TRUE) %>% # Remove duplicates based on ID
  separate(`Login ID`, into = c("Sample", "ID"), sep = 4) %>% # Split ID in Sample and Number
  mutate(
    ID = as.numeric(ID)
    ) %>%
  select(
         Sample, # Sample
         ID, # ID
         employment.what_is_your_highest_level_of_education, #highest level of edcuation
         employment.government_work_key_workers, # key worker status
         employment.government_work_key_workers, # key worker status
         employment.unemployed, #key worker routing variable
         employment.zerohours_contract, #key worker routing variable
         employment.stayathome_parent_or_carer, #key worker routing variable
         employment.retired, #key worker routing variable
         employment.student_gcse_or_a_level, #key worker routing variable
         employment.student_university, #key worker routing variable
         employment.became_unemployed, #key worker routing variable
         employment.furloughed_or_paid_leave_government_funded, #key worker routing variable
         employment.furloughed_or_paid_leave_company_funded, #key worker routing variable
         employment.paid_leave_furloughed #key worker routing variable
         ) %>%
  add_numeric(., exclude = exclude_cols_employment) %>%
  mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
  mutate_if(is.numeric, ~na_if(., -99)) %>%
  mutate_if(is.numeric, ~na_if(., -77)) %>%
  mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
  mutate_if(is.factor, ~na_if(., "Don't know")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

@@

>>>>

# Specify cohort
```{r Specify cohort}
GLAD = FALSE
EDGI = FALSE
NBR = FALSE
RAMP = TRUE
GLAD_and_EDGI = FALSE
```

@@

```{r source file path}
#source raw data directory: data.raw_path
source("../PANCHANGE_raw_path.R")
```

@@

## GLAD and EDGI
### Create data frame 
```{r GLAD and EDGI combined}
if(GLAD_and_EDGI == TRUE)
glad.and.edgi.dem.raw.id <- rbind(glad.dem.raw.id, edgi.dem.raw.id)
```

@@

glad.employment.raw.id <- glad.employment.raw %>%
  drop_na(externalDataReference) %>% # Drop NAs
  distinct(externalDataReference, .keep_all = TRUE) %>% # Remove duplicates based on ID
  separate(externalDataReference, into = c("Sample", "ID"), sep = 4) %>% # Split ID in Sample and Number
  mutate(
    ID = as.numeric(ID)
    ) %>%
  select(
         Sample, # Sample
         ID, # ID
         employment.phd, # level of education - univeristy
         employment.masters_degree_or_equivalent, # level of education - univeristy
         employment.postgraduate_degree_or_equivalent, # level of education - univeristy
         employment.bachelors_degree_or_equivalent, # level of education - univeristy
         employment.a_levelsas_levels_or_equivalent, # level of education - alevel
         employment.o_levelsgcses_or_equivalent, # level of education - gcse
         employment.cses_or_equivalent, # level of education - cse
         employment.nvq_or_hnd_or_hnc_or_equivalent, # level of education - nvq
         employment.government_work_key_workers, # key worker status
         employment.government_work_key_workers, # key worker status
         employment.unemployed, #key worker routing variable
         employment.zerohours_contract, #key worker routing variable
         employment.stayathome_parent_or_carer, #key worker routing variable
         employment.retired, #key worker routing variable
         employment.student_gcse_or_a_level, #key worker routing variable
         employment.student_university, #key worker routing variable
         employment.became_unemployed, #key worker routing variable
         employment.furloughed_or_paid_leave_government_funded, #key worker routing variable
         employment.furloughed_or_paid_leave_company_funded, #key worker routing variable
         employment.paid_leave_furloughed, #key worker routing variable
         employment.fulltime_employed,
         employment.parttime_employed,
         employment.selfemployed,
         employment.contract_or_freelance_work,
         employment.small_business_owner, 
         employment.receiving_state_income,
         employment.my_employment_status_has_not_changed, 
         employment.reduction_in_hours, 
         employment.reduction_in_salary,
         employment.benefits_increased,
         employment.benefits_decreased,
         employment.became_employed,
         employment.became_unemployed,
         employment.increased_hours,
         employment.increased_salary,
                  ) %>%
  add_numeric(., exclude = exclude_cols_employment) %>%
  mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
  mutate_if(is.numeric, ~na_if(., -99)) %>%
  mutate_if(is.numeric, ~na_if(., -77)) %>%
  mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
  mutate_if(is.factor, ~na_if(., "Don't know")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

@@

edgi.employment.raw.id <- edgi.employment.raw %>%
  drop_na(externalDataReference) %>% # Drop NAs
  distinct(externalDataReference, .keep_all = TRUE) %>% # Remove duplicates based on ID
  separate(externalDataReference, into = c("Sample", "ID"), sep = 4) %>% # Split ID in Sample and Number
#  separate("Login ID", into = c("COPING_Sample", "COPING_ID"), sep = 6) %>% # Split ID in Sample and Number
  mutate(
    ID = as.numeric(ID),
#    COPING_ID = as.numeric(COPING_ID)
    ) %>%
  select(
         Sample, # Sample
         ID, # ID
         COPING_ID = "Login ID",
         employment.phd, # level of education - univeristy
         employment.masters_degree_or_equivalent, # level of education - univeristy
         employment.postgraduate_degree_or_equivalent, # level of education - univeristy
         employment.bachelors_degree_or_equivalent, # level of education - univeristy
         employment.a_levelsas_levels_or_equivalent, # level of education - alevel
         employment.o_levelsgcses_or_equivalent, # level of education - gcse
         employment.cses_or_equivalent, # level of education - cse
         employment.nvq_or_hnd_or_hnc_or_equivalent, # level of education - nvq
         employment.government_work_key_workers, # key worker status
         employment.government_work_key_workers, # key worker status
         employment.unemployed, #key worker routing variable
         employment.zerohours_contract, #key worker routing variable
         employment.stayathome_parent_or_carer, #key worker routing variable
         employment.retired, #key worker routing variable
         employment.student_gcse_or_a_level, #key worker routing variable
         employment.student_university, #key worker routing variable
         employment.became_unemployed, #key worker routing variable
         employment.furloughed_or_paid_leave_government_funded, #key worker routing variable
         employment.furloughed_or_paid_leave_company_funded, #key worker routing variable
         employment.paid_leave_furloughed, #key worker routing variable
         employment.fulltime_employed,
         employment.parttime_employed,
         employment.selfemployed,
         employment.contract_or_freelance_work,
         employment.small_business_owner, 
         employment.receiving_state_income,
         employment.my_employment_status_has_not_changed, 
         employment.reduction_in_hours, 
         employment.reduction_in_salary,
         employment.benefits_increased,
         employment.benefits_decreased,
         employment.became_employed,
         employment.became_unemployed,
         employment.increased_hours,
         employment.increased_salary
                  ) %>%
  add_numeric(., exclude = exclude_cols_employment_edgi) %>%
  mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
  mutate_if(is.numeric, ~na_if(., -99)) %>%
  mutate_if(is.numeric, ~na_if(., -77)) %>%
  mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
  mutate_if(is.factor, ~na_if(., "Don't know")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

@@

nbr.employment.raw.id <- nbr.employment.raw %>%
  drop_na(subjectid) %>% # Drop NAs
  distinct(subjectid, .keep_all = TRUE) %>% # Remove duplicates based on ID
  separate(subjectid, into = c("Sample", "ID.intm"), sep = 6) %>% # Split ID in Sample and Number
  separate(ID.intm, into = "ID", sep = 7) %>%
  mutate(
    ID = as.numeric(ID)) %>%
  select(
         Sample, # Sample
         ID, # ID
         employment.phd, # level of education - univeristy
         employment.masters_degree_or_equivalent, # level of education - univeristy
         employment.postgraduate_degree_or_equivalent, # level of education - univeristy
         employment.bachelors_degree_or_equivalent, # level of education - univeristy
         employment.a_levelsas_levels_or_equivalent, # level of education - alevel
         employment.o_levelsgcses_or_equivalent, # level of education - gcse
         employment.cses_or_equivalent, # level of education - cse
         employment.nvq_or_hnd_or_hnc_or_equivalent, # level of education - nvq
         employment.government_work_key_workers, # key worker status
         employment.unemployed, #key worker routing variable
         employment.zerohours_contract, #key worker routing variable
         employment.stayathome_parent_or_carer, #key worker routing variable
         employment.retired, #key worker routing variable
         employment.student_gcse_or_a_level, #key worker routing variable
         employment.student_university, #key worker routing variable
         employment.became_unemployed, #key worker routing variable
         employment.furloughed_or_paid_leave_government_funded, #key worker routing variable
         employment.furloughed_or_paid_leave_company_funded, #key worker routing variable
         employment.paid_leave_furloughed, #key worker routing variable
         employment.fulltime_employed,
         employment.parttime_employed,
         employment.selfemployed,
         employment.contract_or_freelance_work,
         employment.small_business_owner, 
         employment.receiving_state_income,
         employment.my_employment_status_has_not_changed, 
         employment.reduction_in_hours, 
         employment.reduction_in_salary,
         employment.benefits_increased,
         employment.benefits_decreased,
         employment.became_employed,
         employment.became_unemployed,
         employment.increased_hours,
         employment.increased_salary
         ) %>%
  add_numeric(., exclude = exclude_cols_employment) %>%
  mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
  mutate_if(is.numeric, ~na_if(., -99)) %>%
  mutate_if(is.numeric, ~na_if(., -77)) %>%
  mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
  mutate_if(is.factor, ~na_if(., "Don't know")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

@@

ramp.employment.raw.id <- ramp.employment.raw %>%
  drop_na(`Login ID`) %>% # Drop NAs
  distinct(`Login ID`, .keep_all = TRUE) %>% # Remove duplicates based on ID
  separate(`Login ID`, into = c("Sample", "ID"), sep = 4) %>% # Split ID in Sample and Number
  mutate(
    ID = as.numeric(ID)
    ) %>%
  select(
         Sample, # Sample
         ID, # ID
         employment.what_is_your_highest_level_of_education, #highest level of edcuation
         employment.government_work_key_workers, # key worker status
         employment.government_work_key_workers, # key worker status
         employment.unemployed, #key worker routing variable
         employment.zerohours_contract, #key worker routing variable
         employment.stayathome_parent_or_carer, #key worker routing variable
         employment.retired, #key worker routing variable
         employment.student_gcse_or_a_level, #key worker routing variable
         employment.student_university, #key worker routing variable
         employment.became_unemployed, #key worker routing variable
         employment.furloughed_or_paid_leave_government_funded, #key worker routing variable
         employment.furloughed_or_paid_leave_company_funded, #key worker routing variable
         employment.paid_leave_furloughed, #key worker routing variable
         employment.fulltime_employed,
         employment.parttime_employed,
         employment.selfemployed,
         employment.contract_or_freelance_work,
         employment.small_business_owner, 
         employment.receiving_state_income,
         employment.my_employment_status_has_not_changed, 
         employment.reduction_in_hours, 
         employment.reduction_in_salary,
         employment.benefits_increased,
         employment.benefits_decreased,
         employment.became_employed,
         employment.became_unemployed,
         employment.increased_hours,
         employment.increased_salary
                  ) %>%
  add_numeric(., exclude = exclude_cols_employment) %>%
  mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
  mutate_if(is.numeric, ~na_if(., -99)) %>%
  mutate_if(is.numeric, ~na_if(., -77)) %>%
  mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
  mutate_if(is.factor, ~na_if(., "Don't know")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

@@

## Employment prior to pandemic (baseline, GLAD, EDGI, NBR)
Proceed along qualifications fields. If field is not blank, take that as highest qual. If field is blank, check next field. If all fields are blank, NA
```{r Employment prior to pandemic baseline}
if(GLAD == TRUE | EDGI == TRUE | NBR == TRUE) {
#create numeric version of the highest education variable
employment.raw.id <- employment.raw.id %>%
  mutate(
    employment_prior_covid_numeric =
      case_when(
        employment.fulltime_employed_numeric == "1" ~ 1,
        employment.parttime_employed_numeric == "1" ~ 1,
        employment.unemployed_numeric == "1" ~ 2,
        employment.zerohours_contract_numeric == "1" ~ 1,
        employment.selfemployed_numeric == "1" ~ 1,
        employment.contract_or_freelance_work_numeric == "1" ~ 1,
        employment.small_business_owner_numeric == "1" ~ 1,
        employment.retired_numeric  ==  "1" ~ 4,
        employment.student_gcse_or_a_level_numeric == "1" ~ 3,
        employment.student_university_numeric == "1" ~ 3
      )
    )

#recode the numeric version into a factor
employment.raw.id <- employment.raw.id %>%
  mutate(
    employment_prior_covid =
      recode_factor(
        employment_prior_covid_numeric,
        `1` = "Employed",
        `2` = "Unemployed",
        `3` = "Student",
        `4` = "Retired")
    )

employment.raw.id %>%
  freq(employment_prior_covid)
}
```

Check Employment prior to pandemic coding
```{r Employment prior to pandemic check}
if(GLAD == TRUE | EDGI == TRUE | NBR == TRUE) {
employment.raw.id %>%
  select(employment_prior_covid,
         employment_prior_covid_numeric,
        employment.fulltime_employed_numeric, 
        employment.parttime_employed_numeric, 
        employment.unemployed_numeric, 
        employment.zerohours_contract_numeric, 
        employment.selfemployed_numeric,
        employment.contract_or_freelance_work_numeric,
        employment.small_business_owner_numeric,
        employment.retired_numeric,
        employment.student_gcse_or_a_level_numeric,
        employment.student_university_numeric
  )
}
```

## Employment prior to pandemic (baseline, RAMP)
```{r highest level of education for ramp}
if(RAMP == TRUE) {
#create numeric version of the highest education variable
employment.raw.id <- employment.raw.id %>%
  mutate(
    employment_prior_covid_numeric =
      case_when(
        employment.fulltime_employed_numeric == "1" ~ 1,
        employment.parttime_employed_numeric == "1" ~ 1,
        employment.unemployed_numeric == "1" ~ 2,
        employment.zerohours_contract_numeric == "1" ~ 1,
        employment.selfemployed_numeric == "1" ~ 1,
        employment.contract_or_freelance_work_numeric == "1" ~ 1,
        employment.small_business_owner_numeric == "1" ~ 1,
        employment.retired_numeric  ==  "1" ~ 4,
        employment.student_gcse_or_a_level_numeric == "1" ~ 3,
        employment.student_university_numeric == "1" ~ 3
      )
    )

#recode the numeric version into a factor
employment.raw.id <- employment.raw.id %>%
  mutate(
    employment_prior_covid =
      recode_factor(
        employment_prior_covid_numeric,
        `1` = "Employed",
        `2` = "Unemployed",
        `3` = "Student",
        `4` = "Retired")
    )

employment.raw.id %>%
  freq(employment_prior_covid)
}
```

Check Employment prior to pandemic coding for RAMP
```{r Employment prior to pandemic check for RAMP}
if(RAMP == TRUE) {
employment.raw.id %>%
  select(employment_prior_covid,
         employment_prior_covid_numeric,
        employment.fulltime_employed_numeric, 
        employment.parttime_employed_numeric, 
        employment.unemployed_numeric, 
        employment.zerohours_contract_numeric, 
        employment.selfemployed_numeric,
        employment.contract_or_freelance_work_numeric,
        employment.small_business_owner_numeric,
        employment.retired_numeric,
        employment.student_gcse_or_a_level_numeric,
        employment.student_university_numeric
  )
}
```

## Employment change (baseline, GLAD, EDGI, NBR)
Proceed along qualifications fields. If field is not blank, take that as highest qual. If field is blank, check next field. If all fields are blank, NA
```{r Employment change baseline}
if(GLAD == TRUE | EDGI == TRUE | NBR == TRUE) {
#create numeric version of the highest education variable
employment.raw.id <- employment.raw.id %>%
  mutate(
    employment_change_numeric =
      case_when(
        employment.reduction_in_hours_numeric == "1" ~ 1,
        employment.reduction_in_salary_numeric == "1" ~ 1,
        employment.became_unemployed_numeric == "1" ~ 1,
        employment.benefits_decreased_numeric == "1" ~ 1,
        employment.increased_hours_numeric == "1" ~ 2,
        employment.increased_salary_numeric == "1" ~ 3,
        employment.became_employed_numeric == "1" ~ 3,
        employment.benefits_increased_numeric  ==  "1" ~ 2,
        employment.furloughed_or_paid_leave_government_funded_numeric == "1" ~ 3,
        employment.furloughed_or_paid_leave_company_funded_numeric == "1" ~ 3,
        employment.paid_leave_furloughed_numeric == "1" ~ 3,
        employment.my_employment_status_has_not_changed_numeric == "1" ~ 4
      )
    )

#recode the numeric version into a factor
employment.raw.id <- employment.raw.id %>%
  mutate(
    employment_change =
      recode_factor(
        employment_change_numeric,
        `1` = "Decreased employment",
        `2` = "Increased employment",
        `3` = "Furloughed",
        `4` = "Employment not changed")
    )

employment.raw.id %>%
  freq(employment_change)
}
```

Check Employment change coding
```{r Employment change check}
if(GLAD == TRUE | EDGI == TRUE | NBR == TRUE) {
employment.raw.id %>%
  select(employment_change,
         employment_change_numeric,
        employment.reduction_in_hours_numeric,
        employment.reduction_in_salary_numeric,
        employment.became_unemployed_numeric,
        employment.benefits_decreased_numeric,
        employment.increased_hours_numeric, 
        employment.increased_salary_numeric, 
        employment.became_employed_numeric,
        employment.benefits_increased_numeric, 
        employment.furloughed_or_paid_leave_government_funded_numeric,
        employment.furloughed_or_paid_leave_company_funded_numeric, 
        employment.paid_leave_furloughed_numeric, 
        employment.my_employment_status_has_not_changed_numeric
  )
}
```

## Employment change (baseline, RAMP)
```{r employment change for ramp}
if(RAMP == TRUE) {
#create numeric version of the highest education variable
employment.raw.id <- employment.raw.id %>%
  mutate(
    employment_change_numeric =
      case_when(
        employment.reduction_in_hours_numeric == "1" ~ 1,
        employment.reduction_in_salary_numeric == "1" ~ 1,
        employment.became_unemployed_numeric == "1" ~ 1,
        employment.benefits_decreased_numeric == "1" ~ 1,
        employment.increased_hours_numeric == "1" ~ 2,
        employment.increased_salary_numeric == "1" ~ 3,
        employment.became_employed_numeric == "1" ~ 3,
        employment.benefits_increased_numeric  ==  "1" ~ 2,
        employment.furloughed_or_paid_leave_government_funded_numeric == "1" ~ 3,
        employment.furloughed_or_paid_leave_company_funded_numeric== "1" ~ 3,
        employment.paid_leave_furloughed_numeric == "1" ~ 3,
        employment.my_employment_status_has_not_changed_numeric == "1" ~ 4
      )
    )

#recode the numeric version into a factor
employment.raw.id <- employment.raw.id %>%
  mutate(
    employment_change =
      recode_factor(
        employment_change_numeric,
        `1` = "Decreased employment",
        `2` = "Increased employment",
        `3` = "Furloughed",
        `4` = "Employment not changed")
    )

employment.raw.id %>%
  freq(employment_change)
}
```

Check Employment changen coding for RAMP
```{r Employment change check for RAMP}
if(RAMP == TRUE) {
employment.raw.id %>%
  select(employment_change,
         employment_change_numeric,
        employment.reduction_in_hours_numeric,
        employment.reduction_in_salary_numeric,
        employment.became_unemployed_numeric,
        employment.benefits_decreased_numeric,
        employment.increased_hours_numeric, 
        employment.increased_salary_numeric, 
        employment.became_employed_numeric,
        employment.benefits_increased_numeric, 
        employment.furloughed_or_paid_leave_government_funded_numeric,
        employment.furloughed_or_paid_leave_company_funded_numeric, 
        employment.paid_leave_furloughed_numeric, 
        employment.my_employment_status_has_not_changed_numeric
  )
}
```
